---
title: "Lloyd Smith exercises"
author: "Gaurav Kandlikar"
date: "September 16, 2015"
output: html_document
---

### Mini exercise 3.2.1

Tweaking parameters of discrete growth model

```{r}
# conditions & parameters (these are new!)
generations <- 100
RR <- 0.9 # R is a parameter
N_init <- 100 # N is a state variable

# create and populate pop size vector
pop_sizes <- numeric(generations)
pop_sizes[1] <- N_init
for (time in 2:generations){
  pop_sizes[time] <- RR*pop_sizes[time-1]
}

head(pop_sizes)

# My favourite plot:
plot(x = 1:generations, y = pop_sizes, type = "l", lwd = 2,
     xlab = "generations", main = paste("R = ", RR))

```

### Exercise 3.2.2
Convert the model to a function

```{r}
# The function
geom_pop_growth <- function(N, RR, generations = 20) {
  pop_sizes <- numeric(generations)
  pop_sizes[1] <- N_init
  for (time in 2:generations){
    pop_sizes[time] <- RR*pop_sizes[time-1]
  }
  plot(y = pop_sizes, x = 1:generations, type = "l", lwd = 2)
  return(pop_sizes)
}

# Run the function 
geom_pop_growth(N = 100, RR = 1.1)
```


### Exercise 3.3
Logistic growth (includes carrying capacity)

```{r}
# parameters
# conditions & parameters (these are new!)
generations <- 200
rr <- .1 # R is a parameter
N_init <- 100 # N is a state variable
K <- 1000 # Carrying capacity - a parameter

# create and populate pop size vector
pop_sizes <- numeric(generations)
pop_sizes[1] <- N_init
for (time in 2:generations){
  pop_sizes[time] <- pop_sizes[time-1]*(1+rr*(1-pop_sizes[time-1]/K))
}
plot(y = pop_sizes, x = 1:generations, type = "l", lwd = 2)

```

### 3.3.1
Convert above logistic growth to a function

```{r}
# The function

logistic_pop_growth <- function(N, rr, generations = 20, K, plot = T) {
  pop_sizes <- numeric(generations)
  pop_sizes[1] <- N
  for (time in 2:generations){
    pop_sizes[time] <- pop_sizes[time-1]*(1+rr*(1-pop_sizes[time-1]/K))
  }
  
  if(plot == T) {plot(y = pop_sizes, x = 1:generations, type = "l", lwd = 2)}
  return(pop_sizes)
}

# Run the function
logistic_pop_growth(N = 10, rr = 0.1, K = 100)

```

### Exercises 3.3.2

Do a systematic exploration of parameter space

```{r}
rrs <- c(-0.3, 0.3, 1.3, 1.9, 2.2, 2.7)

par(mfrow = c(2,3))
for(rr in rrs) {
  logistic_pop_growth(N = 10, rr = rr, generations = 20, K = 100)
}
```


### Exercise 3.3.3

Bifurcation plot

```{r}
rrs <- seq(from = -1, to = 3, length.out = 100)
generations <- 100
trajectories <- sapply(rrs, function(x) logistic_pop_growth(N = 10, K = 100, rr = x, generations = generations, plot = F))

# I think this is what a bifurcation plot is!
par(mfrow = c(1,1))
matplot(y = trajectories[generations, ], x = rrs, type = "l", lwd = 2, main = "Bifurcation plot?")

# Or maybe it's this, but so messy!
matplot(y = trajectories, x = rrs, type = "l", lwd = 0.5)
```

### 4.1: Differential equations with `deSolve`

The `lsoda` function in `deSolve` is a good numerical differential equation solver. 
```{r}
library(deSolve)

# Set up the model
expGrowthODE <- function(tt, NN, params) { # Note that this needs to be (time, init, params!)
  derivs <- params["rr"]*NN # Exponential growth equation
  return(list(derivs))  # Return the derivs 
}

# Use lsoda to run the ODE
# lsoda(initial value of state variable, time seq, user-defined model, parameters)

# Set initial and parameters
init <- 1
tseq <- seq(from = 1, to = 20, by = 0.01)
parameters <- c(rr = 0.1)

pop_growth <- lsoda(init, tseq, expGrowthODE, parameters)
head(pop_growth)
plot(x = pop_growth[, 1], y = pop_growth[, 2], type = "l", lwd = 2,
     xlab = "time", ylab = "population size")
```

### Exercise 4.2.1
Building the logistic growth model
```{r}
# Set up the model
logGrowthODE <- function(tt, NN, params) { 
  derivs <- params["rr"]*NN*(1-(NN/params["kk"]))
  return(list(derivs))  # Return the derivs 
}

# Set initial and parameters
init <- 10
tseq <- seq(from = 1, to = 125, by = 0.01)
parameters <- c(rr = 0.05, kk = 100)

log_pop_growth <- lsoda(init, tseq, logGrowthODE, parameters)
head(log_pop_growth)
plot(x = log_pop_growth[, 1], y = log_pop_growth[, 2], type = "l", lwd = 2, xlab = "time", ylab = "population size")

# plot dN/dT vs time
log_pop_growth <- data.frame(log_pop_growth)
log_pop_growth[,3] <- rep(NA, nrow(log_pop_growth))
# log_pop_growth[1, 3] <- 0
# Get dN per dt
differences <- diff(log_pop_growth[,2])
log_pop_growth[2:nrow(log_pop_growth), 3] <- differences
plot(x = log_pop_growth[, 1], y = log_pop_growth[, 3], type = "l", lwd = 2,
     xlab = "time", ylab = "change in population size")

# Fix names of log_pop_growth to be a bit more readable
colnames(log_pop_growth)

colnames(log_pop_growth) <- c("time", "pop_size", "growth"); colnames(log_pop_growth)

# make a stacked plot
par(mfrow = c(2, 1))
plot(x = log_pop_growth[, 1], y = log_pop_growth[, 2], type = "l", lwd = 2,
     xlab = "time", ylab = "population size")
plot(x = log_pop_growth[, 1], y = log_pop_growth[, 3], type = "l", lwd = 2,
     xlab = "time", ylab = "change in population size")

# Find out where growth rate (dN/dT) peaks
max_growth <- log_pop_growth[which(log_pop_growth[,3] == max(log_pop_growth[,3], na.rm = T)), ]
max_growth

# Make a single plot, maybe :o
par(mfrow = c(1,1), mar = c(5,4,4,4)+0.3)
plot(x = log_pop_growth[, 1], y = log_pop_growth[, 2], type = "l", lwd = 3, xlab = "time", ylab = "population size", main = "Population with logistic growth")
par(new = T)
plot(x = log_pop_growth[, 1], y = log_pop_growth[, 3], type = "l", lwd = 3, xlab = "", ylab = "", axes = F, col = "darkgreen")
axis(side = 4, at = pretty(range(log_pop_growth[, 3], na.rm = T)))
mtext(side = 4, "growth rate", line = 3)
text(15, .012,  "growth rate", col = "darkgreen")
text(80, .012,  "population size")
abline(v = max_growth[, 1], lwd = 2, lty = 2)
mtext(side = 1, text = paste("Max growth rate occurs when N = ", floor(max_growth[,2]), sep = ""), line = 4)
```

Playing around a bit more with the logistic growth model:

```{r}
# Set initial and parameters
init <- 10
tseq <- seq(from = 1, to = 125, by = 0.01)

# In exercise 3.3.2, we explored values of rr. 
# Let's do that again, but with the continuous version
# Recall that earlier, rrs <- c(-0.3, 0.3, 1.3, 1.9, 2.2, 2.7)
# We can do this in a loop.

rrs <- c(-0.3, 0.3, 1.3, 1.9, 2.2, 2.7) 
par(mfrow = c(2,3))
for (ii in rrs) {
  parameters <- c(rr = ii, kk = 100)
  log_pop_growth2 <- lsoda(init, tseq, logGrowthODE, parameters)
  head(log_pop_growth2)
  plot(y = log_pop_growth2[,2], x = log_pop_growth2[,1], type = "l", lwd = 3, xlab = "time", ylab = "population size", main = paste("r = ", ii))
  }
```

Compare this to the figure in Exercise 3.3.2, where the population sizes bounced around at higher values of R. This may because converting it to continuous time may make the model respond immediately to being close to the carrying capacity as opposed to responding to some previous time step. **look this up!**

### Exercise 4.2.2
How will the dynamics of the logistic model be impacted by harvesting (e.g. by fishing or hunting of the population)? A simple model that incorporated a constant per capita risk of being harvested is coded below:

```{r}
logGrowthHarvestODE <- function(tt, NN, params) { 
  derivs <- params["rr"]*NN*(1-(NN/params["kk"]))-params["hh"]*NN
  return(list(derivs))  # Return the derivs 
}
```

*If you remember how, analyze this model to find its equilibria*  

**Note: still a work in progress, working on this!**
Attempt:
`dN/dT = rN*(1-N/K)*(hN) = 0`  
Equilibrium occurs where dN/dT = 0. This happens in three scenarios (at least, I think?):  
1. N = 0 (trivial)  
2. N = K(1-h/r)
3. h = r(1-N/K)

```{r}
# Set initial and parameters
init <- 10
tseq <- seq(from = 1, to = 125)
parameters <- c(rr = 0.1, kk = 100, hh = 0.01 )
log_pop_growth_harvest <- lsoda(init, tseq, logGrowthHarvestODE, parameters)
head(log_pop_growth_harvest)
plot(y = log_pop_growth_harvest[, 2], x = log_pop_growth_harvest[, 1], main = "logistic growth with harvest", lwd = 2, type = "l", ylab = "pop size", xlab = "time")
```

Assess sensitivity of the model to values of `h`:

```{r}
parameters <- c(rr = 0.1, kk = 100, hh = 0.01 )

par(mfrow = c(2,3))
for (ii in hhs) {
  parameters["hh"] <- ii
  log_pop_growth_harvest <- lsoda(init, tseq, logGrowthHarvestODE, parameters)
  plot(y = log_pop_growth_harvest[, 2], x = log_pop_growth_harvest[, 1], main = "logistic growth with harvest", lwd = 2, type = "l", ylab = "pop size", xlab = "time")

}

```


```{r}
# Write a function to calculate final population size given parameters
sensitiv_harvest <- function(hh){
  parameters["hh"] <- hh
  temp <- lsoda(init, tseq, logGrowthHarvestODE, parameters)
  plot(y = temp[, 2], x = temp[, 1], main = "logistic growth with harvest", lwd = 2, type = "l", ylab = "pop size", xlab = "time")
  return(temp[,2])
}

# make a vector of hhs and run model over it
hhs <- seq(from = 0, to = 0.2, by = 0.005)
sensitivity_to_h <- sapply(hhs, function(x) sensitiv_harvest(x))

plot(y = sensitivity_to_h[nrow(sensitivity_to_h), ], x = hhs, xlab = "h", ylab = "final population size", main = paste("r =", parameters["rr"]), type = "l", lwd = 4)

```
This makes sense - as the proportion of individuals harvested increases, the final population size decreases. 


Algebraicly, dN/dT should be 0 when `h = r(1-N/K)`. Let's confirm that this is true. **NOTE: haven't gotten anywhere with this either. Don't bother for now.**

```{r}
head(sensitivity_to_h)
# Lets write a function again to take a vector, compute differences, and plot.


dNdT_v_h <- function(vec) {
  dNs <- diff(vec)
  x_axis_values <- parameters["rr"]*(1-vec/parameters["kk"])
  x_axis_values <- x_axis_values[2:length(x_axis_values)]
  
  plot(x = x_axis_values, y = dNs)
}

placeholder <- apply(sensitivity_to_h, 1, dNdT_v_h)

```
