---
title: "exercises"
author: "Gaurav Kandlikar"
date: "September 16, 2015"
output: html_document
---

### Exercise 1
Write a for loop statements so that it runs from 1:9 and prints the following output to your screen:
```{r example 1}
for(i in 1:9) {
  if (i < 9) {
    cat("\n")
  } else {
    cat("*")
  }
}
```


### Exercise 2
Modify your for loop so that it prints 10 asterisks, with each asterisk separated by exactly one ampersand sign (&), with no spaces or new line characters.

```{r example 2}
for(i in 1:9) {
  if (i < 9) {
    cat("*&")
  } else {
    cat("*&*")
  }
}

```

### Exercise 3
By hand, figure out the initial values of these variables and values at the the start and end of each iteration of the loop

```{r}
dogs <- 10;
for (i in 1:5){
    dogs <- dogs + 1; 
}

# Initial value of dogs is 10
# final value will be 15 (10 + 1 = 11 + 1 = 12 + 1 = 13 + 1 = 14 + 1 = 15)

dogs
```

```{r}
meatloaf <- 0; 
for (i in 5:9){
    meatloaf <- meatloaf - i + 1;
    # cat(meatloaf) 
}

# initial value of meatloaf is 0
# final value is 0-5+1 = -4; -4-6+1 = -9; -9-7+1 = -15; -15-8+1 = -22; -22-9+1 = -30

meatloaf
```

```{r}
bubbles <- 12;
for (i in -1:-4){
    bubbles <- i;
}

# initial value of bubbles is 12;
# bubbles is then set to -1,then -2, ... -4
bubbles
```

### Exercise 4
Modify this code so that it will print out a message during presidential as well as congressional election years

```{r}
years <- c( 2015, 2016, 2018, 2020, 2021)
for(ii in 1:length(years)){
  if(years[ii] %% 2 == 0){
    cat(years[ii], 'Hooray, congressional elections!', sep = '\t', fill = T)
  } 
  if(years[ii] %% 4 == 0) {
    cat(years[ii], 'Hooray, presidential elections!', sep = '\t', fill = T)
  }
}
```

### Exercise 5 
More fun with loops. Here are the bank accounts from seven randomly selected UCLA grad students

Look at the error message the following lines of code produce. Can you think of a way to modify this loop so that the loop will compund the interest?

```{r}
bankAccounts <- c(10, 9.2, 5.6, 3.7, 8.8, 0.5);


interestRate <- 0.0125
for (i in 1:length(bankAccounts)) {
    compounded[i] <- interestRate*bankAccounts[i] + bankAccounts[i]; }

# we need to initialize `compounded`

compounded <- rep(NA, length(bankAccounts))
for (i in 1:length(bankAccounts)) {
    compounded[i] <- interestRate*bankAccounts[i] + bankAccounts[i]; }
```

### Exercise 6

Go back to the compounded interest example. Suppose we now want to compound the interest annually, but across a period of 5 years. The for loop we discussed earlier only compounds for a single year. Try this:

```{r}
bankAccounts <- c(10, 9.2, 5.6); #define bank accounts here
interestRate <- 0.0525;   
house <- c(4.8, 3.8, 5.7); #deduct
food<- c(3.5, 4.3, 5.0);    #deduct
fun <- c(7.8, 2.1, 10.5);  #deduct
# and incomes (through TAships)
income <- c(21, 21, 21); #add this

for (j in 1:5) {
    for (i in 1:length(bankAccounts)) {
      #step 1 modify bankAccounts so that amounts reflect income and expenses
      bankAccounts[i] <- bankAccounts[i]-house[i]-food[i]-fun[i]+income[i]
      #step 2 get calculate interest and add to accounts from step 1
      bankAccounts[i] <- interestRate*bankAccounts[i] + bankAccounts[i];
      #you can actually use the line you have already written if you
      #modify amounts in bankAccounts directly in step 1
    }
}

bankAccounts
```


### Exercise 7
Three students have estimated annual expenditures for food, housing, and fun of: (in thousands of dollars)
```{r}
bankAccounts <- c(10, 9.2, 5.6) # define bank accounts here
interestRate <- 0.0525   
house <- c(4.8, 3.8, 5.7) # deduct
food <-  c(3.5, 4.3, 5.0) # deduct
fun <-  c(7.8, 2.1, 10.5) # deduct

# and incomes (through TAships) of 

income <- c(21, 21, 21);


for (j in 2015:2020) {
    for (i in 1:length(bankAccounts)) {
      # Update account with deductions and income
      bankAccounts[i] <- bankAccounts[i]-house[i]-food[i]-fun[i]+income[i]
      
      # Update with fellowship money if available
      if (j %% 2 == 1) { # Odd numbers
        if (i == 1 | i == 3) { # For students 1 and 3
          bankAccounts[i] <- bankAccounts[i] + 5
        }
      }
      
      bankAccounts[i] <- interestRate*bankAccounts[i] + bankAccounts[i]
      # you can actually use the line you have already written if you
      # modify amounts in bankAccounts directly in step 1
    }
}
bankAccounts
```


### Exercise 8
Use a while loop to sum all numbers from 1:17. You will need to use a counter variable (like index seen in class).

```{r}
counter <- 1
total = 0

while (counter < 18) {
  total <- total + counter
  counter <- counter + 1
}

# Make sure it's right!
total == sum(1:17)
```

### Exercise 9
Write a function takes a number, and prints ‘small’ if number less than or equal to -1; ‘medium’ if between -1 and + 1’big’ if greater than or equal to + 1

```{r}
size <- function(number) {
  if (number <= -1) {
    print("small")
  } else if (number > -1 & number < 1) {
    print("medium")
  } else {
    print("big")
  }
}

# Try to test this using apply
numbers <- seq(from = -1.2, to = 1.2, by = 0.6)
sapply(numbers, size) # Not sure why this guy returns an extra vector :(

# using a for loop...
for (ii in numbers){
  size(ii)
}
```